---
output: github_document
bibliography: vignettes/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Travis build status](https://travis-ci.org/venelin/PCMBaseCpp.svg?branch=master)](https://travis-ci.org/venelin/PCMBaseCpp)
[![Coverage status](https://codecov.io/gh/venelin/PCMBaseCpp/branch/master/graph/badge.svg)](https://codecov.io/github/venelin/PCMBaseCpp?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/PCMBaseCpp?color=blue)](https://cran.r-project.org/package=PCMBaseCpp)
[![Downloads](http://cranlogs.r-pkg.org/badges/PCMBaseCpp?color=blue)](https://cran.r-project.org/package=PCMBaseCpp)


# PCMBaseCpp

This is a fast C++ backend for the [PCMBase](https://venelin.github.io/PCMBase) R-package. 

# Installation

The package needs a C++ 11 compiler and Rcpp to be installed in you R-environment. Once this is done, you can install the most recent version of the package from github:

```{r, eval=FALSE}
devtools::install_github("venelin/PCMBaseCpp")
```

If you experience problems installing the package from github, you may try installing a possibly older version from CRAN:

```{r, eval = FALSE}
install.packages("PCMBaseCpp")
```

Once the package is installed, use the function `BenchmarkRvsCpp` to evaluate the gain in speed of the likelihood calculation on your machine, relative to the R implementation:

```{r}
library(PCMBaseCpp)
options(digits = 4)

# Depending on your use case, you can change the number of traits, as well as the 
# other arguments:
benchRes <- BenchmarkRvsCpp(ks = 2, includeParallelMode = FALSE, verbose = TRUE)
```

# How to use the package?

There are two ways to use the package:

## Passing the function `PCMInfoCpp` as a `metaI` argument of `PCMLik` and/or `PCMCreateLikelihood`

```{r}
library(PCMBase)
library(PCMBaseCpp)

system.time(llR <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab))

system.time(llCpp <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = PCMInfoCpp))

print(llR)
print(llCpp)
```


```{r}
logLikFunR <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

logLikFunCpp <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = PCMInfoCpp)

randParam <- PCMParamRandomVecParams(PCMBaseTestObjects$model_MixedGaussian_ab)

system.time(llR <- logLikFunR(randParam))

system.time(llCpp <- logLikFunCpp(randParam))

print(llR)
print(llCpp)
```

## Passing the meta-information object returned by `PCMInfoCpp` as a `metaI` argument of `PCMLik` and `PCMCreateLikelihood`

This is the recommended usage in the case of multiple likelihood evaluations, e.g. during model inference:
```{r}
metaIR <- PCMInfo(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

metaICpp <- PCMInfoCpp(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

system.time(llR <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaIR))

system.time(llCpp <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp))

print(llR)
print(llCpp)
```

```{r}
logLikFunR <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaIR)

logLikFunCpp <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp)

system.time(llR <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaIR))

system.time(llCpp <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp))

print(llR)
print(llCpp)
```

# Citing PCMBase

To give credit to the PCMBase package in a publication, please cite the following articles:

Mitov, V., & Stadler, T. (2018). Parallel likelihood calculation for phylogenetic comparative models: The SPLITT C++ library. Methods in Ecology and Evolution, 2041–210X.13136. http://doi.org/10.1111/2041-210X.13136

Mitov, V., Bartoszek, K., Asimomitis, G., & Stadler, T. (2018, September 24). Fast likelihood evaluation for multivariate phylogenetic comparative methods: the PCMBase R package. arXiv.org. https://arxiv.org/abs/1809.09014.

# Used 3rd party libraries

```{r create-references, echo=FALSE, include=FALSE, eval=TRUE}
rcpp <- c("Rcpp")
dataProcessing <- c("data.table")
maths <- c("RcppArmadillo")
testing <- c("testthat", "covr")
docs <- c("roxygen2", "pkgdown")

packagesUsed <- c(rcpp, dataProcessing, maths, testing, docs)

printPackages <- function(packs) {
  res <- ""
  for(i in 1:length(packs)) {
    res <- paste0(res, paste0(packs[i], ' v', packageVersion(packs[i]), ' [@R-', packs[i], ']'))
    if(i < length(packs)) {
      res <- paste0(res, ', ')
    }
  }
  res
}

# Write bib information (this line is executed manually and the bib-file is edited manually after that)
knitr::write_bib(packagesUsed, file = "./vignettes/REFERENCES-R.bib")
```

The PCMBaseCpp R-package uses the following R-packages and C++ libraries:

* For tree processing in C++: The [SPLITT library](https://venelin.github.io/SPLITT) [@Mitov:2018dqa];
* For data processing in R: `r printPackages(dataProcessing)`;
* For algebraic manipulation: The [Armadillo C++ template library](http://arma.sourceforge.net/) [@Sanderson:2016cs] and its port to R `r printPackages(maths)`;
* For unit-testing: `r printPackages(testing)`;
* For documentation and web-site generation: `r printPackages(docs)`;

# Licence and copyright

Copyright 2016-2019 Venelin Mitov

Source code to PCMBaseCpp is made available under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version. PCMBaseCpp is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details.

# References