---
output: github_document
bibliography: vignettes/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(PCMBase)
library(PCMBaseCppCustom)
```

# PCMBaseCppCustom

This is a customized version of the fast C++ backend for the PCMBase R-package. This version adds a C++ implementation for the Early Burst (EB) phylogenetic comparative model.

# Installation

The package needs a C++ 11 or later compiler and Rcpp to be installed in your R-environment. Once this is done, you can install the most recent version of the package from github:

```{r, eval=FALSE}
devtools::install_github("MehediHasanSM/PCMBaseCppCustom", force = TRUE)
```

# Citing PCMBase

To give credit to the PCMBase package in a publication, please cite the following articles:

Mitov, V., & Stadler, T. (2018). Parallel likelihood calculation for phylogenetic comparative models: The SPLITT C++ library. Methods in Ecology and Evolution, 2041–210X.13136. <http://doi.org/10.1111/2041-210X.13136>

Mitov, V., Bartoszek, K., Asimomitis, G., & Stadler, T. (2019). Fast likelihood calculation for multivariate Gaussian phylogenetic models with shifts. Theor. Popul. Biol. <https://doi.org/10.1016/j.tpb.2019.11.005>

# Used 3rd party libraries

```{r create-references, echo=FALSE, include=FALSE, eval=TRUE}
rcpp <- c("Rcpp")
dataProcessing <- c("data.table")
maths <- c("RcppArmadillo")
testing <- c("testthat", "covr")
docs <- c("roxygen2", "pkgdown")

packagesUsed <- c(rcpp, dataProcessing, maths, testing, docs)

printPackages <- function(packs) {
  res <- ""
  for(i in 1:length(packs)) {
    res <- paste0(res, paste0(packs[i], ' v', packageVersion(packs[i]), ' [@R-', packs[i], ']'))
    if(i < length(packs)) {
      res <- paste0(res, ', ')
    }
  }
  res
}

# Write bib information (this line is executed manually and the bib-file is edited manually after that)
knitr::write_bib(packagesUsed, file = "./vignettes/REFERENCES-R.bib")
```

The PCMBaseCppCustom R-package uses the following R-packages and C++ libraries:

-   For tree processing in C++: The [SPLITT library](https://venelin.github.io/SPLITT/) [@Mitov:2018dqa];
-   For data processing in R: `r printPackages(dataProcessing)`;
-   For algebraic manipulation: The [Armadillo C++ template library](https://arma.sourceforge.net/) [@Sanderson:2016cs] and its port to R `r printPackages(maths)`;
-   For unit-testing: `r printPackages(testing)`;
-   For documentation and web-site generation: `r printPackages(docs)`;

# References
