---
title: "Getting started with the PCMBaseCpp R-package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# How to use the package?

There are two ways to use the package:

## Passing the function `PCMInfoCpp` as a `metaI` argument of `PCMLik` and/or `PCMCreateLikelihood`

```{r}
library(PCMBase)
library(PCMBaseCpp)

system.time(llR <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab))

system.time(llCpp <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = PCMInfoCpp))

print(llR)
print(llCpp)
```


```{r}
logLikFunR <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

logLikFunCpp <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = PCMInfoCpp)

metaICpp <- PCMInfoCpp(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

logLikFunCpp2 <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp)

set.seed(1, kind = "Mersenne-Twister", normal.kind = "Inversion")
randParam <- PCMParamRandomVecParams(PCMBaseTestObjects$model_MixedGaussian_ab)

system.time(llR <- logLikFunR(randParam))

system.time(llCpp <- logLikFunCpp(randParam))

system.time(llCpp2 <- logLikFunCpp2(randParam))

print(llR)
print(llCpp)
print(llCpp2)
```

## Passing the meta-information object returned by `PCMInfoCpp` as a `metaI` argument of `PCMLik` and `PCMCreateLikelihood`

This is the recommended usage in the case of multiple likelihood evaluations, e.g. during model inference:
```{r}
metaIR <- PCMInfo(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

metaICpp <- PCMInfoCpp(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab)

system.time(llR <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaIR))

system.time(llCpp <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp))

print(llR)
print(llCpp)
```

```{r}
logLikFunR <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaIR)

logLikFunCpp <- PCMCreateLikelihood(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp)

system.time(llR <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaIR))

system.time(llCpp <- PCMLik(
  X = PCMBaseTestObjects$traits.ab.123, 
  tree = PCMBaseTestObjects$tree.ab,
  model = PCMBaseTestObjects$model_MixedGaussian_ab, 
  metaI = metaICpp))

print(llR)
print(llCpp)
```